Division and Slope Factorization of p-Adic Polynomials

### 2016-02-01 polynomials/muldivfact.tex
typos

The rest of this subsection is devoted to the proof of the theorem.
If $d = n$ (resp. $d = 0$), the sequence $A_i$ is constant equal to
$P$ (resp. to the constant coefficient of $P$) and theorem is clear.
We then assume $0 < d < n$. Without loss of generality, we may further 
assume that the coefficient $a_d$ is equal to $1$. We set:
\begin{align*}
\lambda_0 & = \NF(P)(d) - \NF(P)(d{-}1) \\
\lambda_1 & = \NF(P)(d{+}1) - \NF(P)(d),
\end{align*}
so that $\kappa = \lambda_1 - \lambda_0$.
The existence of an extremal point of $\NP(P)$ located at abscissa
$d$ ensures that $\lambda_1 > \lambda_0$, \emph{i.e.} $\kappa > 0$.
For all indices $i$, we define:
$$\begin{array}{r@{\,\,}lr@{\,\,}l}
Q_i &= V_i P \div A_i, &
R_i & = V_i P \mod A_i = A_{i+1} - A_i, 
\smallskip \\
S_i &= P \mod A_i, &
T_i &= (1 - V_i B_i) \mod A_i
\end{array}$$
and when $\square$ is some letter, we put $\Delta \square_i = \square_{i+1} 
- \square_i$.
%We use the next Lemma to record some useful formulae.

@@

We define the function $\varphi : \R^+ \to \R 
\cup \{+\infty\}$ by:
\begin{equation} \label{eqdef:functionphi}
\varphi(x) = \begin{cases} \NF(P)(x) &\mbox{ if $x \leq d$} \\
\lambda_0 (x - d) + \NF(P)(d) &\mbox{ if $x > d$} \end{cases}
%\begin{array}{r@{\,\,}ll}
%\varphi(x) & = \NP(P)(x) & \text{for } x \leq d \smallskip \\
%& = \lambda_0 (x-d) + \NP(P)(d) & \text{for } x > d.
%\end{array}
\end{equation}
We notice that, when the polynomial is changed into $cP$ where $c$ is a 
nonzero constant lying in a finite extension of $K$, the $A_i$'s are all 
multiplied by $c$ as well whereas the $B_i$'s and the $V_i$'s remained 
unchanged. Therefore, the theorem holds for $P$ if and only if it holds 
for $cP$. As a consequence we may assume that $P$ is normalized so that 
$\NF(P)(d) = d \lambda_0$, so that $\varphi(x) = \lambda_0 x$ for $x > d$.
For a polynomial $Q \in K[X]$ of degree $n$, we further define:
\begin{align}
b_\varphi(Q) & = \min_{x \in [0,n]} \NF(Q)(x) - \varphi(x) 
\label{eqdef:bphi} \\
\text{and} \quad
b_i(Q) & = \min_{x \in [0,n]} \NF(Q)(x) - \lambda_i x
\quad \text{for } i \in \{0,1\}.
\label{eqdef:b0}
\end{align}
Set also $b_\varphi(0) = b_0(0) = b_1(0) = +\infty$ by convention.
With the normalization of $P$ we chose above, we have $b_0(P)
= b_\varphi(P) = 0$ and $b_\varphi(Q) \leq b_0(Q)$ for all polynomial
$Q$. Similarly $b_1(Q) \leq b_0(Q)$ for all $Q$.

@@

\caption{Bound on $\NF(R_0)$ and $\NF(S_0)$}
\label{fig:boundR0S0}
\end{figure}
%
Finally observe that:
$$T_0 = (1 - B_0) \mod A_0 = \big((A_0 - P) \div A_0\big) \mod A_0.$$
Therefore $b_0(T_0) \geq \kappa$ results from $b_\varphi(A{-}P_0) \geq 
\kappa$ thanks to Lemma~\ref{lem:bphib0}. 
We have then established~\eqref{eq:induction} when $i = 0$.

We now assume~\eqref{eq:induction} for the index $i$.
From $A_{i+1} = A_i + R_i$ and the estimation $b_\varphi(R_i) \geq 2^i 
\kappa > 0$, we derive $\NF(A_{i+1}) = \varphi_{|[0,d]}$. Therefore,
Lemma~\ref{lem:bphib0} applies with $A = A_i$ and $A = A_{i+1}$.
Now coming back to the the definition of $B_{i+1}$ and using
Theorem~\ref{theo:EDivisionNP}, we get $b_0(B_{i+1}) \geq b_1(B_{i+1}) 
\geq 0$. As a consequence:
$$b_\varphi(R_i B_{i+1}) \geq b_\varphi(R_i) + b_0(B_{i+1})
\geq 2^i \kappa$$
by Lemma~\ref{lem:bphib0} and the induction hypothesis.
Using again Lemma~\ref{lem:bphib0}, we then derive from 
\eqref{eqdef:Biminus} and \eqref{eqdef:Siminus}, $b_0(\Delta B_i) \geq 
2^i \kappa$ and $b_\varphi(\Delta S_i) \geq 2^i \kappa$.
Similarly, using \eqref{eqdef:Si2} and the estimations we already know,
we obtain $b_\varphi(S_i) \geq 2^i \kappa$. Combining this with
$b_\varphi(\Delta S_i) \geq 2^i \kappa$, we find
$b_\varphi(S_{i+1}) \geq 2^i \kappa$ as well. Applying again and again
the same strategy, we deduce successively
$b_0(\Delta V_i) \geq 2^i \kappa$ using~\eqref{eqdef:Viminus},
$b_0(1{-}Q_i) \geq 2^i \kappa$ using~\eqref{eqdef:Qiminus},
$b_\varphi(R_{i+1}) \geq 2^{i+1} \kappa$ using~\eqref{eq:Riplus},
and then $b_0(T_{i+1}) \geq 2^{i+1} \kappa$
using~\eqref{eq:ViBiplus}. Finally, coming back to the recurrence
defining $V_{i+1}$ and remembering that $b_1(V_i)$ and $b_1(B_{i+1})$
are both nonnegative, we find $b_1(V_{i+1}) \geq 0$.
The equalities and inequalities
of Eq.~\eqref{eq:induction} have then all been established for
the index $i+1$ and the induction goes.

>>>>

The rest of this subsection is devoted to the proof of the theorem.
If $d = n$ (resp. $d = 0$), the sequence $A_i$ is constant equal to
$P$ (resp. to the constant coefficient of $P$) and theorem is clear.
We then assume $0 < d < n$. We set:
\begin{align*}
\lambda_0 & = \NF(P)(d) - \NF(P)(d{-}1) \\
\lambda_1 & = \NF(P)(d{+}1) - \NF(P)(d),
\end{align*}
so that $\kappa = \lambda_1 - \lambda_0$.
The existence of an extremal point of $\NP(P)$ located at abscissa
$d$ ensures that $\lambda_1 > \lambda_0$, \emph{i.e.} $\kappa > 0$.
For all indices $i$, we define:
$$\begin{array}{r@{\,\,}lr@{\,\,}l}
Q_i &= V_i P \div A_i, &
R_i & = V_i P \mod A_i = A_{i+1} - A_i, 
\smallskip \\
S_i &= P \mod A_i, &
T_i &= (1 - V_i B_i) \mod A_i
\end{array}$$
and when $\square$ is some letter, we put $\Delta \square_i = \square_{i+1} 
- \square_i$.
%We use the next Lemma to record some useful formulae.

@@

We define the function $\varphi : \R^+ \to \R 
\cup \{+\infty\}$ by:
\begin{equation} \label{eqdef:functionphi}
\varphi(x) = \begin{cases} \NF(P)(x) &\mbox{ if $x \leq d$} \\
\lambda_0 (x - d) + \NF(P)(d) &\mbox{ if $x > d$} \end{cases}
%\begin{array}{r@{\,\,}ll}
%\varphi(x) & = \NP(P)(x) & \text{for } x \leq d \smallskip \\
%& = \lambda_0 (x-d) + \NP(P)(d) & \text{for } x > d.
%\end{array}
\end{equation}
We notice that, when the polynomial $P$ is changed into $cP$ where $c$ is a 
nonzero constant lying in a finite extension of $K$, the $A_i$'s are all 
multiplied by $c$ as well whereas the $B_i$'s and the $V_i$'s remained 
unchanged. Therefore, the theorem holds for $P$ if and only if it holds 
for $cP$. As a consequence we may assume that $P$ is normalized so that 
$\NF(P)(d) = d \lambda_0$, \emph{i.e.} $\varphi(x) = \lambda_0 x$ for $x > d$.
For a polynomial $Q \in K[X]$ of degree $n$, we further define:
\begin{align}
b_\varphi(Q) & = \min_{x \in [0,n]} \NF(Q)(x) - \varphi(x) 
\label{eqdef:bphi} \\
\text{and} \quad
b_i(Q) & = \min_{x \in [0,n]} \NF(Q)(x) - \lambda_i x
\quad \text{for } i \in \{0,1\}.
\label{eqdef:b0}
\end{align}
Set also $b_\varphi(0) = b_0(0) = b_1(0) = +\infty$ by convention.
With the normalization of $P$ we chose above, we have $b_0(P)
= b_\varphi(P) = 0$ and $b_\varphi(Q) \leq b_0(Q)$ for all polynomial
$Q$. Similarly $b_1(Q) \leq b_0(Q)$ for all $Q$.

@@

\caption{Bound on $\NF(R_0)$ and $\NF(S_0)$}
\label{fig:boundR0S0}
\end{figure}
%
Finally observe that:
$$T_0 = (1 - B_0) \mod A_0 = \big((A_0 - P) \div A_0\big) \mod A_0.$$
Therefore $b_0(T_0) \geq \kappa$ results from $b_\varphi(A_0{-}P) \geq 
\kappa$ thanks to Lemma~\ref{lem:bphib0}. 
We have then established~\eqref{eq:induction} when $i = 0$.

We now assume~\eqref{eq:induction} for the index $i$.
From $A_{i+1} = A_i + R_i$ and the estimation $b_\varphi(R_i) \geq 2^i 
\kappa > 0$, we derive $\NF(A_{i+1}) = \varphi_{|[0,d]}$. Therefore,
Lemma~\ref{lem:bphib0} applies with $A = A_i$ and $A = A_{i+1}$.
Now coming back to the the definition of $B_{i+1}$ and using
Theorem~\ref{theo:EDivisionNP}, we get $b_0(B_{i+1}) \geq b_1(B_{i+1}) 
\geq 0$. As a consequence:
$$b_\varphi(R_i B_{i+1}) \geq b_\varphi(R_i) + b_0(B_{i+1})
\geq 2^i \kappa$$
by Lemma~\ref{lem:bphib0} and the induction hypothesis.
Using again Lemma~\ref{lem:bphib0}, we then derive from 
\eqref{eqdef:Biminus} and \eqref{eqdef:Siminus} that $b_0(\Delta B_i) \geq 
2^i \kappa$ and $b_\varphi(\Delta S_i) \geq 2^i \kappa$.
Similarly, using \eqref{eqdef:Si2} and the estimations we already know,
we obtain $b_\varphi(S_i) \geq 2^i \kappa$. Combining this with
$b_\varphi(\Delta S_i) \geq 2^i \kappa$, we find
$b_\varphi(S_{i+1}) \geq 2^i \kappa$ as well. Applying again and again
the same strategy, we deduce successively
$b_0(\Delta V_i) \geq 2^i \kappa$ using~\eqref{eqdef:Viminus},
$b_0(1{-}Q_i) \geq 2^i \kappa$ using~\eqref{eqdef:Qiminus},
$b_\varphi(R_{i+1}) \geq 2^{i+1} \kappa$ using~\eqref{eq:Riplus},
and then $b_0(T_{i+1}) \geq 2^{i+1} \kappa$
using~\eqref{eq:ViBiplus}. Finally, coming back to the recurrence
defining $V_{i+1}$ and remembering that $b_1(V_i)$ and $b_1(B_{i+1})$
are both nonnegative, we find $b_1(V_{i+1}) \geq 0$.
The equalities and inequalities
of Eq.~\eqref{eq:induction} have then all been established for
the index $i+1$ and the induction goes.
